---
title: "Using a Shared Drive"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)

```

## Intro

In the Mayor's Office of Recovery Programs, we call it the "shared drive" but I think it's more commonly called a "Network Drive". These are folders that can be setup and shared within an office or agency. They aren't on an individual's computer (they're on a server) but they look like a native folder after they're [mapped](https://support.microsoft.com/en-us/windows/map-a-network-drive-in-windows-29ce55d1-34e3-a7e2-4801-131475f9557d).

These are a few examples using a Windows computer that's on the City's network. The examples read and write straight to the network drive; no files are saved locally. The examples use excel files but this process will work with any file.


## Pros and Cons

### Pros

+ Data is securely saved on a server where others can access
+ Reading and writing feels familiar
+ other can access it, including non-data people

### Cons

+ Need to be on the city network
+ Might be limited to Windows machines?
+ No Audit trail or backups



## Examples

### Reading files

```{r}
### libraries
library(tidyverse)



### Read a file straight from the network drive

## path to directory on our agency's shared drive
projects_dir <- "\\\\balt-smb-nas\\MORP\\Funding_Announcement\\"

## list files in this directory of the network drive
list.files(projects_dir)

#add the file to the end of path
file_to_read <- paste0(projects_dir, "Funding_Announcement.xlsx")

## this would print the whole path, including file
print(file_to_read)

### since it's an excel file, using this function 
### to read it from the network drive

readxl::read_xlsx(file_to_read) %>% 
select(1:4)

```


### Writing

```{r}

### This will write the mtcars dataframe straight to a file in the network drive

## destintation on drive where file will go, including filename
file_destination <- paste0(projects_dir, "mtcars.xlsx")

# show the whole path including file name
file_destination

# write it directly on networkdrive
writexl::write_xlsx(mtcars, file_destination)

```


 