---
title: "Using documents in sharepoint"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)

```

## Intro

Reading and writing files to SharePoint offer a few advantages over local files and even network drives. Using SharePoint documents for data storage also comes with more complicated functions, workflow, and setup. First, interacting with SharePoint requires the Graph API. Setting up the graph api requires registration with BCIT. An overview of the registration and examples using R can be found [in this repo](https://github.com/city-of-baltimore/graph-api-example/tree/main/R/microsoft365r).

The following examples uses a a graph API app setup using the `client_credentials` flow, which I would generally recommend. 


## Pros and Cons

### Pros

+ Data is securely saved in sharepoint where others can access, including non-data people
+ You can see who edited files in sharepoint
+ You can rollback to a previous version of a file if something happens
+ The graph api credentials can be used widely thus data can be read by shiny apps and more.
+ Don't need to be on city network

### Cons

+ There's a process to get registered with Graph API
+ You have to manage secrets
+ Additional libraries and functions are required to learn


## Examples

### Connect to Sharepoint

The following code would work better in a few functions, but is presented line-by-line for the example. The example uses excel files for data. Excel opens nicely in the sharePoint UI but this process works with any file type including csv, txt, rds, feather, and other files that might be better for data storage.

```{r}

library(tidyverse)
library(Microsoft365R)
library(AzureGraph)


### establish a connection that will be used for reading and writing
### credentials are stored as environmental variables so no exposed in the code

conn <- AzureGraph::create_graph_login(tenant = "bmore", 
                                       app = Sys.getenv("graph_id"), 
                                       password = Sys.getenv("graph_secret"))
```





### Reading files saved in SharePoint documents

In the example, the file is temporarily downloaded from sharepoint, then read into memory before being deleted. It uses the connection established above. 

```{r}
### File that will be read
file_name <- "prebudget_project_commitments.xlsx"

## whole path to file in SharePoint
whole_file_path <- paste0("Data resources/dashboard_extra_data/", 
                     file_name)

## method to get to sharepoint document functions
### this example uses the MORP SP site, change that to your own
sp_drive <- conn$get_sharepoint_site(site_url = "https://bmore.sharepoint.com/sites/MORP")$get_drive()
  
  
## download the file temporarily 
sp_drive$get_item(whole_file_path)$download()

### read the downloaded file into memory
df <- readxl::read_xlsx(file_name)

## delete file
unlink(file_name)


### show some data
head(df)


```


### Writing a file to Sharepoint

There's a similar process for saving data to sharepoint documents
```{r}

### File that will be written
file_to_write_name <- "mtcars.xlsx"

### Save data to local file
writexl::write_xlsx(mtcars, file_to_write_name)

### upload local file to directory in sharepoint
  sp_drive$upload_file(file_to_write_name,
                          dest = paste0("Data resources/", 
                                        file_to_write_name))

### delete local file
unlink(file_to_write_name)


```



 