---
title: "Using Microsoft SQL Server"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE)

```

## R Markdown

Databases are obviously the standard for remote data storage, though they're a whole different level of security issues, IT, networking, cost, and management. Rstudio has [nice documentation](https://db.rstudio.com/) on best practices for interacting with databases but here are a few code chunks to get started.


## Examples

## Connecting to a SQL database

This can sometimes be the most frustrating part of the whole process. The driver might need to be installed on your machine or server. The R function `odbcListDrivers()` will list any drivers available on your machine, but you might need to install one that's compatible with your database. The driver listed in the code chunk below is free, works with Microsoft SQL server, and I personally have used it on Mac and linux. It's available at [Microsoft](https://docs.microsoft.com/en-us/sql/connect/odbc/download-odbc-driver-for-sql-server?view=sql-server-ver15), but the appropriate driver depends on the type and version of sql server and your OS. **The driver required for shinyapps.io is `SQLServer`**; there are numerous threads that discuss other drivers but this will yield the best results as off July 2022. Also remember that shinyapps.io and many other cloud solutions exist outside the city's network, so you'll need an external IP address and you'll probably need to whitelist the IP addresses of shinyapps.io or the external IP of your R session (your house or azure VM, for example).

The server info is best saved as an environmental variable, and it can come in the form of an IP address or a url-like string. You'll need a user ID and password as well. Depending on where and how the database is hosted, you might also need to be an the network.

```{r}
# libraries
library(tidyverse)
library(odbc)
library(DBI)


conn <- dbConnect( odbc::odbc(),
                        driver = "SQL Server", ## or whatever driver you have installed
                        Server = Sys.getenv("lme_ip"),
                        Database = "lme",
                        Uid = Sys.getenv("lme_id"),
                        Pwd = Sys.getenv("lme_pw") )
```


### Reading Data from a SQL database

```{r}


### writing a query using standard sql, 
### this results in a dataframe with three columns filtered on revenue code

  the_df <- dbGetQuery(conn,
                          "
                          SELECT id, effective_date, status, revenue_code_id
                          FROM quotes
                          WHERE revenue_code_id = 'DYLTL'
                          ")


### writing a query using dplyr syntax, the result is the same
### this is less flexible than writing SQL but more familiar for R users

  the_df <- tbl(conn, 'quotes') %>% 
    filter(revenue_code_id == 'DYLTL') %>% 
    collect()



```



Writing and updating tables

```{r eval=FALSE}
### write a dataframe to a SQL table, this will replace existing table
dbWriteTable(conn, "mtcars_db", mtcars, overwrite = TRUE)


### append a dataframe on to a SQL table
dbWriteTable(conn, "mtcars_db", 
             mtcars, append = TRUE, overwrite =FALSE)


### updating a table. there are a few ways but here's one that works
dbExecute(conn, "UPDATE test_table_update 
                    SET somefield = 'new value'  
                    WHERE name = 'fred'")


```
